---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# finddataanalysis

<!-- badges: start -->
<!-- badges: end -->

The goal of finddataanalysis is to have a collection of tools that are used for data analysis.

## Installation

You can install the development version of finddataanalysis from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("finddx/finddataanalysis")
```


## List of functions

List of functions are included in the help. Run the following code in Console: 
```
help(package ="finddataanalysis")
```





## Performance Evaluations

This is a basic example which shows you how to calculate the sensitivity and specificity of a diagnostic test using the automatized function

```{r example}
library(finddataanalysis)
library(DT)
df <- data.frame(Index1 = c(rep("Positive", 20), rep("Negative", 20)), Index2 = c(rep("Positive", 5), rep("Negative", 35)), Index3 = c(rep("Positive", 9), rep("Negative", 31)), Reference = c(rep("Positive", 10), rep("Negative", 30)))
```


### Generating All outputs
```{r}
eval_output <- performance_eval_auto(data_var = df, list_index = c("Index1", "Index2", "Index3"), ref = "Reference", conf.level = 0.95, index_names = c("Tgs1", "AFD", "SimpleDx"), labels = "Test", forest_plot = TRUE, table_output = TRUE, file_name = "MyEvaluationExample")
eval_output$sen_plot 
eval_output$spe_plot
eval_output$table
```

### Generating Forest Plot Outputs
```{r}
eval_output_only_forest <- performance_eval_auto(data_var = df, list_index = c("Index1", "Index2", "Index3"), ref = "Reference", conf.level = 0.95, index_names = c("Tgs1", "AFD", "SimpleDx"), labels = "Test", forest_plot = TRUE, table_output = FALSE)
eval_output_only_forest$sen_plot
eval_output_only_forest$spe_plot
eval_output_only_forest$table  # Should give NULL
```

### Generating a simple data frame output where the table is not formatted, especially considering the confidence intervals. This form is computer friendly.
```{r}
eval_output_simple_df <- performance_eval_auto(data_var = df, list_index = c("Index1", "Index2", "Index3"), ref = "Reference", conf.level = 0.95, index_names = c("Tgs1", "AFD", "SimpleDx"), labels = "Test", forest_plot = FALSE, table_output = FALSE)

eval_output_simple_df
```

### Performance Evaluation by Groups

You can do the subgroup analysis automatically by defining a group variable (*group_var*). Please note that, this parameter should not be specified in quotation parks.
E.g.: **variable** instead of **"variable"**
```{r}
data(my_dataset)
head(my_dataset)
eval_output <- performance_eval_auto(data_var = my_dataset, list_index = "Result", ref = "RefTest", conf.level = 0.95, labels = "Test", forest_plot = FALSE, table_output = TRUE, file_name = "MyEvaluationExample", data_long = TRUE, group_var = Test_Name )
```

