---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# finddataanalysis

<!-- badges: start -->
<!-- badges: end -->

The goal of finddataanalysis is to have a collection of tools that are used for data analysis.

## Installation

You can install the development version of finddataanalysis from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("berkosarFIND/finddataanalysis")
```


## Performance analysis

This is a basic example which shows you how to calculate the sensitivity and specificity of a diagnostic test.

```{r example}
library(finddataanalysis)
library(DT)
df <- data.frame(Index = c(rep("Positive", 20), rep("Negative", 20)), Reference = c(rep("Positive", 10), rep("Negative", 30)))
dtable <- sens_spe(data_var = df, index = "Index", ref = "Reference", conf.level = 0.95)
DT::datatable(dtable)
```

Here is a forest plot that you can generate:


```{r pressure, echo = FALSE}
library(finddataanalysis)
library(forestplot)
library(dplyr)


df <- data.frame(Index1 = c(rep("Positive", 20), rep("Negative", 20)), Index2 = c(rep("Positive", 15), rep("Negative", 25)),Reference = c(rep("Positive", 30), rep("Negative", 10)))

forest_data1 <- cbind(Test = "Index1", sens_spe_for_forest(data_var = df, index = "Index1", ref = "Reference", conf.level = 0.95))

forest_data2 <- cbind(Test = "Index2", sens_spe_for_forest(data_var = df, index = "Index2", ref = "Reference", conf.level = 0.95))


forest_data <- rbind(forest_data1, forest_data2)
forest_data %>% forestplot(mean = Sensitivity, 
                          lower = SensLower, 
                          upper = SensUpper, 
                          labeltext = Test,
                          title = "Sensitivity",
                          zero = NA,
                          xticks = c( 0, 20, 40, 60, 80, 100),
                          txt_gp = fpTxtGp(ticks=gpar(cex=1)),
                          boxsize = .1
                          )

```
