---
title: "flat_composite_reference.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

<!--
 You need to run the 'description' chunk in the '0-dev_history.Rmd' file before continuing your code there.
-->

# Composite reference

```{r function-composite_reference}
#' composite_reference 
#' 
#' Generates a composite reference based on 3 tests. When any of the tests is positive, subject is classified as positive.
#' Works with 2 or 3 reference tests.
#'
#' @param ref_1 Test1 containing "Positive or "Negative" Results
#' @param ref_2 Test2 containing "Positive or "Negative" Results
#' @param ref_3 Test3 containing "Positive or "Negative" Results
#' @param data_frame  Data table containing the 3 tests
#'
#' @return Data table with composite reference added
#' @export
#' @importFrom plyr rename
#' @import dplyr
#'
#' @examples
composite_reference <- function(ref_1, ref_2, ref_3 = NULL, data_frame){
  if (is.null(ref_3)){
    colname <- paste0(ref_1, "_", ref_2)
    data_frame <- data_frame %>% mutate(col = case_when(data_frame[,ref_1] == "Positive" | data_frame[,ref_2] == "Positive"~ "Positive",
                                                                data_frame[,ref_1] == "Negative" & data_frame[,ref_2] == "Negative"~ "Negative"))

  } else {
    #browser()
    colname <- paste0(ref_1, "_", ref_2, "_", ref_3)
    data_frame <- data_frame %>% mutate( col = case_when(data_frame[,ref_1] == "Positive" | data_frame[,ref_2] == "Positive" | data_frame[,ref_3] == "Positive"~ "Positive",
                                                                data_frame[,ref_1] == "Negative" & data_frame[,ref_2] == "Negative" & data_frame[,ref_3] == "Negative"~ "Negative"))
  }
  data_frame <- plyr::rename(data_frame,c("col"= colname))
  return(data_frame)
}

```

```{r examples-composite}
df <- data.frame(T1 = c(rep("Positive", 10), rep("Negative", 10), rep("Negative", 10)), T2 = c(rep("Positive", 10), rep("Positive", 10), rep("Negative", 10)))
composite_reference(ref_1 = "T1", ref_2 = "T2", ref_3 = NULL, data_frame = df)
```

```{r tests-composite}

```



## Categorise test result based on a threhold
    
```{r function-categorize_result}
#' categorize_result
#' 
#' Categorize a subject as "Positive" or "Negative", based on a variable of interest and a specified threshold. The values are bigger or equal to the threshold are categorized as "Positive", others "Negative".
#' A list of thresholds can be provided.
#' 
#' @param dataframe The data table contaiing test results
#' @param variable A continuous variable
#' @param thresholds The thresholds to be applied
#'
#' @return Data table with the categorization added
#' 
#' @export
categorize_result <- function(dataframe, variable, thresholds){
 for(i in thresholds){
   #browser()
   dataframe <- dataframe %>% mutate (col = ifelse(as.numeric(dataframe[,variable]) >=i, "Positive", "Negative" ))
   dataframe[,"col"]<-  replace(dataframe[,"col"], is.na(dataframe[,"col"]), "Negative")
   dataframe <- plyr::rename(dataframe,c("col"= paste0("Result", i)))
 }
 return(dataframe)
}
```
  
```{r example-categorize_result}

df <- data.frame(results = c(sample(1:100, 50, replace=TRUE)))
categorize_result(dataframe = df, variable = "results", thresholds = 50)
```
  
```{r tests-categorize_result}

```
  
## Categorise test result based on a threhold for PCR (bigger numbers are associated with negativity)
    
```{r function-categorize_result_PCR}
#' categorize_result_PCR
#' 
#' Same as categorize_result but the values are smaller than the threshold are categorized as "Positive", others "Negative".
#' 
#' @param dataframe The data table containing test results
#' @param variable A continuous variable - PCR Ct data
#' @param thresholds The thresholds to be applied
#'
#' @return Data table with category added
#' 
#' @export
categorize_result_PCR <- function(dataframe, variable, thresholds){
  for(i in thresholds){
    #browser()
    dataframe <- dataframe %>% mutate (col = ifelse(as.numeric(dataframe[,variable]) <i, "Positive", "Negative" ))
    dataframe[,"col"]<-  replace(dataframe[,"col"], is.na(dataframe[,"col"]), "Negative")
    dataframe <- plyr::rename(dataframe,c("col"= paste0("Result", i)))
  }
  return(dataframe)
}
```
  
```{r example-categorize_result_PCR}

```
  
```{r tests-categorize_result_PCR}

```


## Composite Reference for ROC

Generates composite references based on 2/3 variables, one being tested for multiple thresholds.
    
```{r function-composite_for_roc}
#' composite_for_roc
#' 
#' Define multiple composite references to make a ROC style analysis but on the references.
#' 
#' @param ref_1 Reference 1
#' @param ref_2  Reference 2
#' @param test_refs Reference categorized based on multiple thresholds
#' @param data_frame Data table
#'
#' @return Data table containing all possible composite references based on the thresholds applied
#' 
#' @export
composite_for_roc <- function(ref_1, ref_2, test_refs, data_frame){
  composite_ref <- data_frame
  for (i in test_refs){
    composite_ref <- composite_reference(ref_1 = ref_1, ref_2 = ref_2, ref_3 = i, data_frame = composite_ref)
  }
  return(composite_ref)
}
```
  
```{r example-composite_for_roc}
```
  
```{r tests-composite_for_roc}

```
  
## Calculate ROC-like caharcterisitcs

This function can be used to monitor sensitivity and specificity of an RDT in comparison to a reference that will be classified based on multiple thresholds.
This is useful in the case where a composite reference is generated based on multiple tests and one of the test results needs to be classified as positive or negative.
The index test classification is not affected.
    
```{r function-ROC_calculate}
#' Title
#' 
#' Calculates sensitivity and false positivity rate based on the multiple reference columns defined by the thresholds
#' 
#' @param data_var data table
#' @param index index test column containing results: "Positive" or "Negative"
#' @param ref a list containing names of the reference test columns
#' @param conf.level confidence level
#' @param var_name_suffix_cut the suffix based on which the reference test column name to be cut. This will result in suffix + threshold
#'
#' @return
#' 
#' @export
ROC_calculate <- function(data_var, index, ref, conf.level = 0.95, var_name_suffix_cut){
  ROC_df <- data_frame()
  name = c()
  for ( i in ref){
    name <- append(name, strsplit(i, var_name_suffix_cut)[[1]][2])
    sen_spe_df <- sens_spe_for_forest(data_var = data_var, index = index, ref = i, conf.level = 0.95)
    ROC_df <- rbind(ROC_df, sen_spe_df)
  }
ROC_df$Ref_Threshold <-  name
ROC_df$FPR <- 100 - ROC_df$Specificity
return(ROC_df)
}
```
  
```{r example-ROC_calculate}

```
  
```{r tests-ROC_calculate}

```
  
## composite_reference_majority
    
```{r function-composite_reference_majority}
#' Composite Reference Majority 
#' 
#' The function reads the results of 3 reference test results and takes the majority as the final result.
#' 
#' #' @param ref_1 Test1 containing "Positive or "Negative" Results
#' @param ref_2 Test2 containing "Positive or "Negative" Results
#' @param ref_3 Test3 containing "Positive or "Negative" Results
#' @param data_frame  Data table containing the 3 tests
#' 
#' 
#' @return Data table with composite reference added
#' 
#' @import dplyr
#' 
#' 
#' @export
composite_reference_majority <- function(ref_1, ref_2, ref_3, data_frame){

    data_frame <- data_frame %>% mutate( MajorityRef = case_when(data_frame[,ref_1] == "Positive" & data_frame[,ref_3] == "Positive" ~ "Positive",
                                                         data_frame[,ref_2] == "Positive" & data_frame[,ref_3] == "Positive" ~ "Positive",
                                                         data_frame[,ref_1] == "Positive" & data_frame[,ref_2] == "Positive" ~ "Positive",
                                                         data_frame[,ref_1] == "Negative" & data_frame[,ref_3] == "Negative" ~ "Negative",
                                                         data_frame[,ref_2] == "Negative" & data_frame[,ref_3] == "Negative" ~ "Negative",
                                                         data_frame[,ref_1] == "Negative" & data_frame[,ref_3] == "Negative" ~ "Negative"))

  return(data_frame)
}
    

```
  
```{r example-composite_reference_majority}
df <- data.frame(T1 = c(rep("Positive", 10), rep("Negative", 10), rep("Negative", 10)), T2 = c(rep("Positive", 10), rep("Positive", 10), rep("Negative", 10)), T3 = c(rep("Negative", 10), rep("Positive", 10), rep("Negative", 10)))
composite_reference_majority(ref_1 = "T1", ref_2 = "T2", ref_3 = "T3", data_frame = df)
```
  
```{r tests-composite_reference_majority}

```
  
```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_composite_reference.Rmd", vignette_name = "composite_reference")
```
